---
title: "Lab05"
author: "Sylvia Baeyens"
date: "9/24/2021"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo= FALSE, include= FALSE}
#including necessary libraries
library(data.table)
library(dplyr)
library(tidyverse)
library(leaflet)
library(R.utils)
library(ggplot2)
```

Loading Data Files & Merging
```{r cache= TRUE}
# Dowloading met data set
if (!file.exists("../met_all.gz")) {
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "../met_all.gz", method="libcurl", timeout = 60)}
met <- data.table::fread("../met_all.gz")

# Download the data
stations <- data.table::fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]

# Dealing with NAs and 999999
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]

# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])

# Dropping NAs
stations <- stations[!is.na(USAF)]

# Removing duplicates
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]

# Merging
met <- merge(
  x = met,
  y = stations,
  all.x = TRUE, all.y = FALSE,
  by.x = "USAFID", by.y = "USAF"
)
```

# 1. Representative Station for the US

First, we find mean values for variables of interest for each station.
```{r}
stationAvg <- met[,.(
  temp = mean(temp, na.rm = TRUE),
  wind.sp = mean(wind.sp, na.rm = TRUE),
  atm.press = mean(atm.press, na.rm = TRUE)
), by = USAFID]
```

Identifying quantiles for variables of interest
```{r}
medians = stationAvg[,.(
  temp50 = quantile(temp, probs= .5, na.rm= TRUE),
  wind50 = quantile(wind.sp, probs= .5, na.rm= TRUE),
  atm50 = quantile(atm.press, probs= .5, na.rm= TRUE)
)]
```

Finding station closest to mean
```{r}
stationAvg[, temp_dist := abs(temp- medians$temp50)]
medianTempStation = stationAvg[order(temp_dist)][1]

stationAvg[, wind_dist := abs(wind.sp- medians$wind50)]
medianWindStation = stationAvg[order(wind_dist)][1]

stationAvg[, atm_dist := abs(atm.press- medians$atm50)]
medianAtmStation = stationAvg[order(atm_dist)][1]
```

The station that is best representative of the median temperature of all stations is `r medianTempStation$USAFID`
The station that is best representative of the median wind speed of all stations is `r medianWindStation$USAFID`
The station that is best representative of the median atmospheric pressure of all stations is `r medianAtmStation$USAFID`

# 2. Representative Station per State

First, we recover state variable via MERGE
```{r}
stationStates = merge(
  x = stationAvg, y = stations,
  by.x = "USAFID", by.y = "USAF",
  all.x = TRUE, all.y = FALSE
  )
```

Now, computing medians per state & eucledian distance
```{r}
stationStates[, temp50 := quantile(temp, probs = .5, na.rm = TRUE), by = STATE]
stationStates[, wind50 := quantile(wind.sp, probs = .5, na.rm = TRUE), by = STATE]
stationStates[, atm50 := quantile(atm.press, probs = .5, na.rm = TRUE), by = STATE]

stationStates[, eudist := sqrt(
  (temp - temp50)^2 + (wind.sp - wind50)^2
  )]
stationStates

medianStateStation = stationStates[order(STATE, eudist)]
```

You can now go through medianStateStation data table to find the most representative per state. For example, in CA, it is station 722970.

# 3. In the middle?


